---
title: "Scenario tutorials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scenario tutorials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(statsguider)
```

This tutorial creates several small data tables and applies `statsguider` to each one.

## Scenario 1: two independent continuous groups

```{r}
tbl_continuous <- data.frame(
  group = c(rep("control", 6), rep("treated", 6)),
  biomarker = c(10.2, 10.4, 10.1, 10.5, 10.3, 10.0, 11.1, 11.4, 11.0, 11.3, 11.5, 11.2)
)

tbl_continuous
```

```{r}
select_test(
  data = tbl_continuous,
  outcome = "biomarker",
  group = "group",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  run = "recommend",
  language = "en"
)
```

```{r}
run_test(
  data = tbl_continuous,
  outcome = "biomarker",
  group = "group",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  language = "en"
)
```

## Scenario 2: paired before-after measurements

```{r}
tbl_paired <- data.frame(
  id = rep(paste0("S", 1:6), each = 2),
  visit = rep(c("before", "after"), times = 6),
  value = c(8.2, 9.0, 7.9, 8.8, 8.5, 9.1, 8.1, 8.9, 8.3, 9.0, 8.0, 8.7)
)

tbl_paired
```

```{r}
select_test(
  data = tbl_paired,
  outcome = "value",
  group = "visit",
  id = "id",
  outcome_type = "continuous",
  paired = "yes",
  repeated = "no",
  run = "recommend",
  language = "en"
)
```

## Scenario 3: three independent continuous groups

```{r}
tbl_three_groups <- data.frame(
  group = rep(c("control", "drugA", "drugB"), each = 5),
  biomarker = c(
    9.8, 10.1, 10.0, 9.9, 10.2,
    10.8, 10.9, 11.1, 10.7, 11.0,
    11.4, 11.2, 11.5, 11.3, 11.6
  )
)

tbl_three_groups
```

```{r}
select_test(
  data = tbl_three_groups,
  outcome = "biomarker",
  group = "group",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  run = "recommend",
  language = "en"
)
```

## Scenario 4: small categorical table

```{r}
tbl_small_cat <- data.frame(
  group = c(rep("control", 6), rep("treated", 6)),
  response = c("yes", "no", "no", "no", "no", "no", "yes", "yes", "no", "no", "no", "no")
)

tbl_small_cat
```

```{r}
select_test(
  data = tbl_small_cat,
  outcome = "response",
  group = "group",
  outcome_type = "binary",
  paired = "no",
  repeated = "no",
  run = "recommend",
  language = "en"
)
```

## Scenario 5: paired binary response

```{r}
tbl_paired_binary <- data.frame(
  id = rep(paste0("P", 1:8), each = 2),
  visit = rep(c("before", "after"), times = 8),
  response = c("no", "yes", "no", "no", "yes", "yes", "no", "yes",
               "yes", "yes", "no", "yes", "no", "no", "yes", "yes")
)

tbl_paired_binary
```

```{r}
select_test(
  data = tbl_paired_binary,
  outcome = "response",
  group = "visit",
  id = "id",
  outcome_type = "binary",
  paired = "yes",
  repeated = "no",
  run = "recommend",
  language = "en"
)
```

## Scenario 6: repeated ordinal data

```{r}
tbl_repeated_ordinal <- data.frame(
  id = rep(paste0("S", 1:5), each = 3),
  visit = rep(c("baseline", "week2", "week4"), times = 5),
  score = ordered(
    c(
      "high", "medium", "low",
      "medium", "medium", "low",
      "high", "medium", "medium",
      "medium", "low", "low",
      "high", "medium", "low"
    ),
    levels = c("low", "medium", "high")
  )
)

tbl_repeated_ordinal
```

```{r}
select_test(
  data = tbl_repeated_ordinal,
  outcome = "score",
  group = "visit",
  id = "id",
  outcome_type = "ordinal",
  paired = "no",
  repeated = "yes",
  run = "recommend",
  language = "en"
)
```

## Scenario 7: stop and redirect because adjustment is needed

```{r}
tbl_adjust <- data.frame(
  group = c(rep("control", 6), rep("treated", 6)),
  age = c(41, 45, 39, 48, 42, 44, 37, 40, 38, 43, 39, 41),
  biomarker = c(10.2, 10.4, 10.1, 10.5, 10.3, 10.0, 11.1, 11.4, 11.0, 11.3, 11.5, 11.2)
)

tbl_adjust
```

```{r}
select_test(
  data = tbl_adjust,
  outcome = "biomarker",
  group = "group",
  outcome_type = "continuous",
  adjust = "yes",
  run = "recommend",
  language = "en"
)
```

## Scenario 8: stop and redirect because the outcome is a count

```{r}
tbl_count <- data.frame(
  group = c(rep("control", 6), rep("treated", 6)),
  colonies = c(12, 10, 11, 14, 9, 13, 18, 17, 20, 16, 19, 18)
)

tbl_count
```

```{r}
select_test(
  data = tbl_count,
  outcome = "colonies",
  group = "group",
  outcome_type = "count",
  paired = "no",
  repeated = "no",
  run = "recommend",
  language = "en"
)
```
