---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(statsguider)
```

`statsguider` is built for a narrow question:

> "I already have a `data.frame`. Which simple statistical test matches this
> design, and can I run it immediately?"

The package keeps the input contract small:

- `data`
- `outcome`
- `group`
- `id` when paired or repeated data are used
- a few choice-style arguments such as `goal`, `outcome_type`, `paired`, and
  `run`

## First recommendation

Create a small table and ask for a recommendation.

```{r}
dat <- data.frame(
  group = c(rep("control", 8), rep("treated", 8)),
  value = c(5.0, 5.2, 4.8, 5.1, 5.3, 4.9, 5.0, 5.4,
            6.1, 6.0, 5.8, 6.2, 6.3, 5.9, 6.1, 6.0)
)

select_test(
  data = dat,
  outcome = "value",
  group = "group",
  goal = "difference",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  adjust = "no",
  normality = "auto",
  run = "recommend"
)
```

## Run the test

When the branch is acceptable, switch to `run = "run"`.

```{r}
select_test(
  data = dat,
  outcome = "value",
  group = "group",
  goal = "difference",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  adjust = "no",
  normality = "auto",
  run = "run"
)
```

## Validate before execution

Use `check_design()` when you want to inspect the design contract explicitly.

```{r}
check_design(
  data = dat,
  outcome = "value",
  group = "group",
  goal = "difference",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  adjust = "no"
)
```

## What happens outside the simple branch

The package redirects rather than forcing a wrong analysis.

```{r}
recommend_test(
  data = dat,
  outcome = "value",
  group = "group",
  goal = "adjusted_effect",
  outcome_type = "continuous",
  paired = "no",
  repeated = "no",
  adjust = "yes"
)
```
