---
title: "Guide: Main Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guide: Main Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(statsguider)
```

This guide summarizes the exported functions.

## `select_test()`

The shortest entry point. It accepts the full argument contract and either
returns a recommendation or runs the selected method.

## `guided_test()`

The interactive entry point. It asks the branching questions step by step. In a
script, pass a named `answers` list instead of answering interactively.

```{r}
dat <- data.frame(
  id = rep(1:6, each = 2),
  group = rep(c("before", "after"), times = 6),
  score = c(4, 6, 5, 7, 6, 8, 7, 8, 5, 6, 4, 5)
)

guided_test(
  data = dat,
  answers = list(
    goal = "difference",
    outcome = "score",
    group = "group",
    paired = "yes",
    repeated = "no",
    adjust = "no",
    id = "id",
    outcome_type = "continuous",
    normality = "no"
  ),
  run = "recommend"
)
```

## `recommend_test()`

Use this when you want a structured branch decision without execution.

## `run_test()`

Use this when you already know the branch fits the package scope and want the
base-R result immediately.

## `check_design()`

Use this when you want explicit validation and warning collection before either
recommendation or execution.
